---
title: "Independent Project"
author: "Brandon Bonilla"
date: "5/2/2022"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "../Output/")

```

## Introduction

The aim of my independent project was to apply what I have learned in class to areas of data analysis that interest me. Sports is a field that heavily relies on data. Data is especially prominent in the case of the Olympics, which incorporates different countries competing throughout history. Many people take pride in the amount of success their country, or an individual athlete see at this stage. We will be taking a look at the two most historically dominant countries in the Olympics, and compare them in the sports they succeed in most. In addition, we will compare male and female athletes of these countries to see if there is any edge between them. 

## Load Libraries

```{r}
library(tidyverse)
library(here)
library(kableExtra)
library(ymlthis)
library(dplyr)
library(patchwork)
library(ggplot2)
library(prettydoc)
```

## Load Data
```{r}
olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')

View(olympics)
```

## Data Analysis: USA & USSR

```{r}
GoldEverything<-olympics%>% # Isolating gold medalists
  filter(medal == "Gold")
view(GoldEverything)

USSRGold<-olympics%>%
  filter(medal == "Gold",
         team == "Soviet Union")

USAGoldSummer<-olympics%>% # Isolating USA gold medalists in basketball 
  filter(season == "Summer",
         medal == "Gold",
         team == "United States",
         sport == "Basketball",
         year >= "1984")%>% # Isolating consecutive golden years
  select(name, sex, team, year, games, sport, event, medal)%>%
  rename("Name" = name,
         "Sex"=sex,
         "Team"=team,
         "Games" =games,
         "Sport"=sport,
         "Event"=event,
         "Medal"=medal,
         "Year"=year)
view(USAGoldSummer)

USAGoldSummerD<-USAGoldSummer%>% # Structure data in descending order by year
  arrange(desc(Year))

USSRGoldSummer<-olympics%>% # Isolating USSR gold medalists in gymnastics
  filter(season == "Summer",
         medal == "Gold",
         team == "Soviet Union",
         sport == "Gymnastics",
         year<= "1980")%>% # Isolating consecutive golden years
  select(name, sex, team, year, games, sport, event, medal)%>%
  rename("Name" = name,
         "Sex"=sex,
         "Team"=team,
         "Games" =games,
         "Sport"=sport,
         "Event"=event,
         "Medal"=medal,
         "Year"=year)
view(USSRGoldSummer)

USSRGoldSummerD<-USSRGoldSummer%>% # Structure data in descending order by year
  arrange(desc(Year))

```

## Table 1. Team USA Basketball
```{r}
Table1<-USAGoldSummerD%>% # Creating table
  group_by(Name, Sex)%>%
  kbl(caption = "Team USA Basketball Gold Medalists")%>%
  kable_material_dark(c("striped"), 
                full_width = FALSE, fixed_thead = T, position = "center", 
                font_size = 12) %>% # Organizing layout for the prettydoc
    row_spec(0, bold = T, color = "black", background = "blue")%>%
  row_spec(1:179, color = "black", background = "grey")
 Table1

```

## Figure 1. Team USA Best Era
```{r, USAplot,fig.align='center'}
USAplot<-USAGoldSummerD%>% # Plotting gold medalists by year
  ggplot( aes(x = Year,
            y = Medal))+
  geom_dotplot(fill = 'gold')+ # dot plot counts each medal
  labs(title = "USA Dominance in Basketball")+
  theme(plot.title = element_text(hjust = 0.5),
        panel.border = element_rect(fill = NA,
                                    color = "blue"))
USAplot
```

## Table 2. Team USSR Gymnastics
```{r}
Table2<-USSRGoldSummerD%>% # Creating table
  group_by(Name, Sex)%>%
  kbl(caption = "Team Soviet Union Gymastics Gold Medalists")%>%
  kable_material_dark(c("striped"), 
                full_width = FALSE, fixed_thead = T, position = "center", 
                font_size = 12) %>% # Organizing layout for the prettydoc
  row_spec(0, bold = T, color = "black", background = "red")%>%
  row_spec(1:120, color ="black", background = "grey")
Table2
```

## Figure 2. Team USSR Best Era
```{r,USSRplot, fig.align='center'}
USSRplot<-USSRGoldSummerD%>% # Plotting gold medalists by year
  ggplot( aes(x = Year,
            y = Medal))+
  geom_dotplot(fill = 'gold')+ # dot plot counts each medal
  labs(title = "USSR Dominance in Gymnastics")+
  theme(plot.title = element_text(hjust = 0.5),
        panel.border = element_rect(fill = NA,
                                    color = "red"))
USSRplot
```

## Figure 3. Comparisson of Dominant Eras
```{r, CompareP, fig.align='center'}
CompareP<-USAplot+USSRplot+ # Patchworking the two figures together
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = 'A')

CompareP
```

## Data Analysis: USA Sexes
```{r}
USAMen<-olympics%>% # USA gold medalists by gender
  filter(season == "Summer",
         sex=="M",
         medal == "Gold",
         team == "United States",
         sport == "Basketball",
         year >= "1984")%>%
  select(name, sex, team, year, games, sport, event, medal)%>%
  rename("Name" = name,
         "Sex"=sex,
         "Team"=team,
         "Games" =games,
         "Sport"=sport,
         "Event"=event,
         "Medal"=medal,
         "Year"=year)
view(USAMen)

USAWomen<-olympics%>% # USA gold medalists by gender
  filter(season == "Summer",
         sex=="F",
         medal == "Gold",
         team == "United States",
         sport == "Basketball",
         year >= "1984")%>%
  select(name, sex, team, year, games, sport, event, medal)%>%
  rename("Name" = name,
         "Sex"=sex,
         "Team"=team,
         "Games" =games,
         "Sport"=sport,
         "Event"=event,
         "Medal"=medal,
         "Year"=year)
view(USAWomen)
```

## Figure 4. USA Men vs Women
```{r, mvf, fig.align='center'}
USAmenp<-USAMen%>% # Plotting by gender
  ggplot( aes(x = Year,
            y = Medal))+
  geom_dotplot(fill = 'gold')+ # dot plot counts each medal
  labs(title = "USA Men's Basketball Winners")+
  theme(plot.title = element_text(hjust = 0.5),
        panel.border = element_rect(fill = NA,
                                    color = "blue"))
USAwomenp<-USAWomen%>% # Plotting by gender
  ggplot( aes(x = Year,
            y = Medal))+
  geom_dotplot(fill = 'gold')+ # dot plot counts each medal
  labs(title = "USA Women's Basketball Winners")+
  theme(plot.title = element_text(hjust = 0.5),
        panel.border = element_rect(fill = NA,
                                    color = "pink"))
mvf<-USAmenp+USAwomenp+ # Patchworking gender figures together
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = 'A')

mvf
```

## Data Analysis USSR Sexes
```{r}
USSRMen<-olympics%>% # USSR gold medalists by gender
  filter(season == "Summer",
         sex == "M",
         medal == "Gold",
         team == "Soviet Union",
         sport == "Gymnastics",
         year<= "1980")%>%
  select(name, sex, team, year, games, sport, event, medal)%>%
  rename("Name" = name,
         "Sex"=sex,
         "Team"=team,
         "Games" =games,
         "Sport"=sport,
         "Event"=event,
         "Medal"=medal,
         "Year"=year)
view(USSRMen)

USSRWomen<-olympics%>% # USSR gold medalists by gender
  filter(season == "Summer",
         sex == "F",
         medal == "Gold",
         team == "Soviet Union",
         sport == "Gymnastics",
         year<= "1980")%>%
  select(name, sex, team, year, games, sport, event, medal)%>%
  rename("Name" = name,
         "Sex"=sex,
         "Team"=team,
         "Games" =games,
         "Sport"=sport,
         "Event"=event,
         "Medal"=medal,
         "Year"=year)
view(USSRWomen)
```

## Figure 5. USSR Men vs. Women
```{r, mvfUSSR, fig.align='center'}
USSRMP<-USSRMen%>% # Plotting by gender
  ggplot( aes(x = Year,
            y = Medal))+
  geom_dotplot(fill = 'gold')+ # dot plot counts each medal
  labs(title = "USSR Men's Gymnastics Winners")+
  theme(plot.title = element_text(hjust = 0.5),
        panel.border = element_rect(fill = NA,
                                    color = "blue"))
USSRWP<-USSRWomen%>% # Plotting by gender
  ggplot( aes(x = Year,
            y = Medal))+
  geom_dotplot(fill = 'gold')+ # dot plot counts each medal
  labs(title = "USSR Women's Gymnastics Winners")+
  theme(plot.title = element_text(hjust = 0.5),
        panel.border = element_rect(fill = NA,
                                    color = "pink"))
mvfUSSR<-USSRMP+USSRWP+
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = 'A')

mvfUSSR
```
### Extra Functions
```{r}

table(USAGoldSummerD['Sex']) # Counting USA gold medals by sex

table(USSRGold['sport']) # Counting USSR gold medals by sport

table(GoldEverything['team']) # Counting all gold medals by team
```


